Evidence for a Color Parameter Unassociated With Dust Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2017-03-29 apj.tex
more simulations for model 11

@@

There is a pair of parameter values,  $k_0^0$ and $k_1^0$, that correspond to supernovae experiencing no extrinsic effect.
The true extrinsic
extinction is
$A^T_X(k_0,k_1)= A_X(k_0,k_1) -  A_X(k^0_0,k^0_1)$ where $A_X(k_0,k_1) =  \gamma^0_X k_0 + \gamma^1_X k_1$.
The true extrinsic color excess is $E^T(B-V) = A^T_B-A^T_V$, so that the total-to-selective extinction is $R^T_V=A^T_V/E^T(B-V)$.
Our $\langle k_0 \rangle=0$ and  $\langle k_1 \rangle=0$ constraints
do not determine $k_0^0$ and $k_1^0$, and so the extrinsic extinction and color excess per supernova are not products of our analysis.

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.95in]{output11/plane0.pdf}
   \includegraphics[width=2.95in]{output11/plane1.pdf}
   \includegraphics[width=2.95in]{output11/plane0BVR.pdf}
   \includegraphics[width=2.95in]{output11/plane1BVR.pdf}
   \caption{
   Two perspectives (left and right columns) of $UVI$- (top row) and $BVR$- (bottom row) spaces
of the vectors corresponding to $\gamma^i_X/(\gamma^i_B-\gamma^i_V)$ of our model (solid lines),
   and $a(X)$, $b(X)$ of the \citetalias{1999PASP..111...63F} model (dashed lines).  The combinations $a(X)+b(X)/2.4$ and  $a(X)+b(X)/2.6$ are shown in the dotted
   lines: they are almost perfectly superimposed on $\gamma^0_X/(\gamma^0_B-\gamma^0_V)$ for $UVI$ and $BVR$ respectively.  $b(X)$ shown in the dotted line is
   almost  perfectly superimposed on $\gamma^1_X/(\gamma^1_B-\gamma^1_V)$ in $UVI$ but is clearly distinct in $BVR$.  
   \label{plane:fig}}
\end{figure}

@@

The preferred values for $\kappa$ are shown qualitatively in
Figure~\ref{plane:fig}.  For $\kappa_1=1/2.4$, $\kappa_2=-6.8$
the vector $\gamma^0_X$ is  aligned with 
the dust vector $a(X)+\kappa_1 b(X)$  and $\gamma^1_X$
is aligned with $\kappa_2 a(X) + b(X)$  $UVI$-space.  
%With these values of $\kappa$, the dust parameters are $R^F_0=2.4$ and $\Delta R^F=1.4$.
Slightly shifted values of $\kappa_1=1/2.6$, $\kappa_2=-6.1$ apply in $BVR$-space.
%, which correspond to  $R^F_0=2.6$ and $\Delta R^F=1.4$.

@@

As a check of the above interpretation, we generate a simulated dataset with the same numbers and uncertainties as our data, based on a
SALT2 $x_1=0$, $c=0$ template and  \citetalias{1999PASP..111...63F} dust.  The supernovae
are assigned $E^F(B-V)$ parameters equal to the $B-V$ color excesses determined in our sample, and $R^F$ values drawn from a log-normal
distribution with a mean of $\ln{2.4}$ and 0.05 standard deviation.  The $\alpha$, $\beta$, and $\eta$ parameters are set to zero, so that dust and
measurement uncertainty are the only sources of variation in magnitudes.  As with the analysis of our data, we find that the
$\gamma$'s and \citetalias{1999PASP..111...63F}  vectors for nearly parallel planes, with the $\gamma^0$ vector almost consistent with
$\kappa_1^{-1}=2.4$, as expected given the mean of $\ln{2.4}$ of the input log-normal distribution.

@@

The resulting credible intervals of our parameters are given in Table~\ref{global2:tab}.  Comparison with the
results of the 2-parameter extrinsic model given in Table~\ref{global:tab} shows that  parameters that are common to both
are consistent well within 1$\sigma$ uncertainties.  The most significant change is in the slight reduction of the standard
deviations of the magnitude residuals $\sigma_X$.
The newly fitted $\gamma$ parameters have an almost identical association with the  \citetalias{1999PASP..111...63F} dust model, in that now
\begin{equation}
\begin{pmatrix}
 \frac{\gamma^0_X}{\gamma^0_1-\gamma^0_2} \\
\frac{\gamma^1_X}{\gamma^1_1-\gamma^1_2} 
\end{pmatrix}=
\begin{pmatrix}
\begin{array}{rr}
1.15 & 2.79 \\
0.71 & -4.95
\end{array}
\end{pmatrix} 
\begin{pmatrix}
a(X) \\
b(X)
\end{pmatrix}+
\begin{pmatrix}
\vec{\epsilon}_a \\
\vec{\epsilon}_b
\end{pmatrix},
\end{equation}
where the residual vectors 
\color{red}
\begin{align}
\vec{\epsilon}_a &=\{0.05, -0.06, -0.06,  0.06,  0.07\},\\
\vec{\epsilon}_b &=\{ -0.10, 0.20,  -0.19,  0.32, -0.30\}
\end{align}
\color{black}
are perpendicular to the  \citetalias{1999PASP..111...63F} plane and
contribute similar 2\% and 4\% to the total  length of the vector.
(These can be compared to the results of Model I in Eqs.\ \ref{trans_I:eqn}--\ref{res_I:eqn}.)
%The effective total-to-selective extinction for the supernova sample has a slight shift to $R^F_V=2.23 \pm 0.16$.

>>>>

\color{orange}
For brevity, we do not present here the analysis of simulated data used to validate and bias quantify bias in our results.
The simulation tests in \S\ref{model2:sec} are representative for both Model~1 of this Section and Model~2 of the subsequent section.
\color{black}

@@

There is a pair of parameter values,  $k_0^0$ and $k_1^0$, that correspond to supernovae experiencing no extrinsic effect.
\color{orange}
For observer filters $X$ and $Y$, we define $A_X(k_0,k_1) =  \gamma^0_X k_0 + \gamma^1_X k_1$, $E(X-Y)=A_X-A_Y$.
The true extrinsic
extinction is
$A^T_X(k_0,k_1)= A_X(k_0,k_1) -  A_X(k^0_0,k^0_1)$.
\color{blue}
The true extrinsic color excess is $E^T(B-V) = A^T_B-A^T_V$, so that the total-to-selective extinction is $R^T_V=A^T_V/E^T(B-V)$.
Our $\langle k_0 \rangle=0$ and  $\langle k_1 \rangle=0$ constraints
do not determine $k_0^0$ and $k_1^0$, and so the extrinsic extinction and color excess per supernova are not products of our analysis.

@@

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=2.95in]{output11/plane0.pdf}
   \includegraphics[width=2.95in]{output11/plane1.pdf}
%   \includegraphics[width=2.95in]{output11/plane0BVR.pdf}
%   \includegraphics[width=2.95in]{output11/plane1BVR.pdf}
   \caption{
   \color{orange}
   Two perspectives (left and right columns) of the $UVI$-space
of the vectors corresponding to $\gamma^i_X/(\gamma^i_B-\gamma^i_V)$ of our model (solid lines),
   and $a(X)$, $b(X)$ of the \citetalias{1999PASP..111...63F} model (dashed lines).  The combination $a(X)+b(X)/2.44$ is shown in the dotted red
   line: it is almost perfectly superimposed on $\gamma^0_X/(\gamma^0_B-\gamma^0_V)$.   The combination $-7.31a(X)+b(X)$ is shown in the dotted blue
   line: it is
   almost  perfectly superimposed on $\gamma^1_X/(\gamma^1_B-\gamma^1_V)$ in $UVI$.  
   \label{plane:fig}}
\end{figure}

@@

\color{orange}
The preferred values for $\kappa^{-1}$ can be determined from the matrix elements of Eqn.~\ref{model11trans:eqn}: $\kappa^{-1}_1=2.82/1.15 =2.44$, $\kappa^{-1}_2=-5.27/0.72=-7.31$.
The directions of the $\gamma$ vectors, and  the expectations from  \citetalias{1999PASP..111...63F} with $R=\kappa^{-1}$, are shown in 
Figure~\ref{plane:fig}.  There is almost perfect alignment between $\gamma^0$ and $R=2.44$ dust.
 From the Monte Carlo chains, we find a 68\% credible interval for $\kappa^{-1}_1$ of 2.26--2.63.
\color{blue}

@@

As a check of the above interpretation, we generate a simulated dataset with the same numbers and uncertainties as our data, based on a
SALT2 $x_1=0$, $c=0$ template and  \citetalias{1999PASP..111...63F} dust.  The supernovae
are assigned $E^F(B-V)$ parameters equal to the $B-V$ color excesses determined in our sample, and $R^F$ values drawn from a log-normal
distribution with a mean of $\ln{2.4}$ and 0.05 standard deviation.  The $\alpha$, $\beta$, and $\eta$ parameters are set to zero, so that dust and
measurement uncertainty are the only sources of variation in magnitudes.  
From the Monte Carlo chains, we find a 68\% credible interval for $\kappa^{-1}_1$ of 2.18--2.40.
As with the analysis of our data, we find that the
$\gamma$'s and \citetalias{1999PASP..111...63F}  vectors for nearly parallel planes, with the $\gamma^0$ vector almost consistent with
$\kappa_1^{-1}=2.4$, as expected given the mean of $\ln{2.4}$ of the input log-normal distribution.

@@

The resulting credible intervals of our parameters are given in Table~\ref{global2:tab}.  Comparison with the
results of the 2-parameter extrinsic model given in Table~\ref{global:tab} shows that  parameters that are common to both
are consistent well within 1$\sigma$ uncertainties.  The most significant change is in the slight reduction of the standard
deviations of the magnitude residuals $\sigma_X$.
The newly fitted $\gamma$ parameters have an almost identical association with the  \citetalias{1999PASP..111...63F} dust model, in that now
\begin{equation}
\begin{pmatrix}
 \frac{\gamma^0_X}{\gamma^0_1-\gamma^0_2} \\
\frac{\gamma^1_X}{\gamma^1_1-\gamma^1_2} 
\end{pmatrix}=
\begin{pmatrix}
\begin{array}{rr}
1.15 & 2.79 \\
0.71 & -4.95
\end{array}
\end{pmatrix} 
\begin{pmatrix}
a(X) \\
b(X)
\end{pmatrix}+
\begin{pmatrix}
\vec{\epsilon}_a \\
\vec{\epsilon}_b
\end{pmatrix},
\end{equation}
where the residual vectors 
\color{red}
\begin{align}
\vec{\epsilon}_a &=\{0.05, -0.06, -0.06,  0.06,  0.07\},\\
\vec{\epsilon}_b &=\{ -0.10, 0.20,  -0.19,  0.32, -0.30\}
\end{align}
\label{model11trans:eqn}
\color{black}
are perpendicular to the  \citetalias{1999PASP..111...63F} plane and
contribute similar 2\% and 4\% to the total  length of the vector.
(These can be compared to the results of Model I in Eqs.\ \ref{trans_I:eqn}--\ref{res_I:eqn}.)
%The effective total-to-selective extinction for the supernova sample has a slight shift to $R^F_V=2.23 \pm 0.16$.
